---
title: "Exploring data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exploring data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.5,
  dpi = 96
)
library(modelviz)
```

Before modelling, it pays to look at the data. modelviz provides four
exploratory plots that share the package theme and palette. Each accepts column
names either quoted or unquoted.

## Distributions

`distribution_plot()` draws a histogram, a density, or both, optionally split
by a group.

```{r}
distribution_plot(lexical_decision, RT, title = "Reaction times")
```

```{r}
distribution_plot(lexical_decision, RT, group = condition, type = "density",
                  title = "Reaction times by priming condition")
```

## Scatter plots with a trend

`scatter_trend_plot()` adds a fitted trend and confidence band; group to fit
one trend per level.

```{r}
scatter_trend_plot(crop_yield, fertilizer, yield, group = treatment,
                   title = "Yield against fertiliser, by treatment")
```

## Correlation heatmaps

`correlation_matrix_plot()` correlates the numeric columns and annotates each
cell. Switch to a rank correlation with `method = "spearman"`.

```{r, fig.height = 5}
correlation_matrix_plot(wellbeing_survey,
                        title = "Correlations among wellbeing measures")
```

## Missing-data maps

`missingness_plot()` shades the missing cells, orders the variables by their
proportion of missing values, and reports the overall rate -- a quick triage
before modelling.

```{r, fig.height = 5}
missingness_plot(wellbeing_survey, title = "Missing values in the survey")
```

The `wellbeing_survey` data were simulated so that income is missing more often
at higher stress (a "missing at random" mechanism), which is the kind of
pattern this plot is designed to reveal.
