---
title: "Getting started with depictr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with depictr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: ../inst/REFERENCES.bib
csl: apa.csl
link-citations: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      fig.width = 7, fig.height = 4.5, dpi = 96)
```

**depictr** is a single, consistent toolkit of plots that span the whole
analysis workflow, from a first look at the data, through model estimates and
predictions, to diagnostics, uncertainty and reporting. Every plotting function
returns a `ggplot2` object [@wickham2016] (or a `patchwork` for composite
panels), so you can keep customising with the usual `+` syntax, and every plot
shares one theme, one palette and one set of label conventions.

```{r}
library(depictr)
```

The package ships with three reproducibly simulated datasets used throughout the
documentation: `lexical_decision`, `wellbeing_survey` and `crop_yield`.

## A tour by task

Begin with the data. `explore_bivariate()` chooses a suitable plot for any pair
of variables.

```{r}
explore_bivariate(crop_yield, fertilizer, yield)
```

Turn next to the model. After fitting it, `coefficient_plot()` draws a forest
plot of the estimates.

```{r}
fit <- lm(yield ~ rainfall + fertilizer + soil_ph + treatment, data = crop_yield)
coefficient_plot(fit, order = "descending", title = "Drivers of crop yield")
```

To see what the model implies, `effects_plot()` traces the predicted response as
one predictor varies.

```{r}
effects_plot(fit, "fertilizer")
```

`residual_diagnostics_plot()` gathers the usual checks of the fit into one
panel.

```{r, fig.height = 6}
residual_diagnostics_plot(fit)
```

Finally, `posterior_plot()` summarises posterior or simulation draws as a point
with nested intervals.

```{r}
set.seed(1)
draws <- data.frame(intercept = rnorm(2000, 2, 0.2),
                    fertilizer = rnorm(2000, 0.012, 0.002),
                    rainfall = rnorm(2000, 0.004, 0.001))
posterior_plot(draws)
```

## The shared spine: `tidy_estimates()`

Most of the model functions rest on `tidy_estimates()`, which turns a model, or
a data frame of pre-computed estimates, into one standard table. Because the
plotting functions also accept that table, estimates from any source (Bayesian
posteriors, bootstrap intervals, or figures taken from a paper) can be supplied
directly.

```{r}
tidy_estimates(fit)
```

## A consistent, accessible look

`theme_depictr()`, `depictr_palette()` and `scale_colour_depictr()` style your
own plots too:

```{r}
library(ggplot2)
ggplot(crop_yield, aes(fertilizer, yield, colour = treatment)) +
  geom_point(alpha = 0.7) +
  scale_colour_depictr() +
  theme_depictr()
```

The qualitative palette is based on the Okabe-Ito set [@okabe2008], which stays
distinguishable under the common forms of colour-vision deficiency; sequential
and diverging variants are available too. Preview them with:

```{r, fig.height = 5}
palette_preview(type = "all")
```

## Where to next

The remaining articles go into each area in turn.
`vignette("exploring-data")` covers distributions, categories, bivariate plots,
scatter-plot matrices, correlations, missingness, outliers and summary tables.
`vignette("model-estimates")` covers forest plots, model comparison, predicted
values, interactions, random effects and goodness-of-fit.
`vignette("diagnostics-and-uncertainty")` covers residuals, influence, Q-Q, ROC,
calibration, confusion matrices, posteriors and power curves. Two further
articles, `vignette("multivariate-and-survival")` and `vignette("time-series")`,
cover the remaining methods.

## References
