---
title: "Getting started with depictr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with depictr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: ../inst/REFERENCES.bib
csl: apa.csl
link-citations: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      fig.width = 7, fig.height = 4.5, dpi = 96)
```

**depictr** is a single, consistent toolkit of plots that span the whole
analysis workflow, from a first look at the data, through model estimates and
predictions, to diagnostics, uncertainty and reporting. Every plotting function
returns a `ggplot2` object [@wickham2016] (or a `patchwork` for composite
panels), so you can keep customising with the usual `+` syntax, and every plot
shares one theme, one palette and one set of label conventions.

```{r}
library(depictr)
```

## Five datasets to explore

The package ships with five reproducibly simulated datasets, each chosen to
exercise a different family of plots. They are documented under their names
(e.g. `?lexical_decision`) and load with `data()`:

* `lexical_decision`: a counterbalanced, crossed reaction-time/accuracy
  experiment (participant, item, condition, modality, word frequency). For
  mixed models and the classification plots.
* `wellbeing_survey`: a cross-sectional survey (life satisfaction, stress,
  sleep, income, age, ordered education, region) with *informative*
  missingness. For descriptives, correlations, regression and missing data.
* `crop_yield`: a field trial with a genuine fertiliser-by-treatment
  interaction. For regression, scatter-trend and interaction plots.
* `clinical_trial`: a two-arm trial with separating survival curves and a
  rare adverse-event outcome. For survival and imbalanced classification.
* `monthly_sales`: two seasonal monthly series (indoor/outdoor). For the
  time-series plots.

## A tour by task

Begin with the data. `explore_bivariate()` chooses a suitable plot for any pair
of variables, here a scatter with a trend because both are numeric.

```{r}
explore_bivariate(crop_yield, fertiliser, yield)
```

Turn next to the model. After fitting it, `coefficient_plot()` draws a forest
plot of the estimates.

```{r}
fit <- lm(yield ~ rainfall + fertiliser + soil_ph + treatment, data = crop_yield)
coefficient_plot(fit, order = "descending", title = "Drivers of crop yield")
```

To see what the model implies, `effects_plot()` traces the predicted response as
one predictor varies.

```{r}
effects_plot(fit, "fertiliser")
```

`residual_diagnostics_plot()` gathers the usual checks of the fit into one
panel.

```{r, fig.height = 6}
residual_diagnostics_plot(fit)
```

For uncertainty, `posterior_plot()` summarises posterior or simulation draws as
a distribution per parameter. These are the real fixed-effect posterior draws
from a Bayesian fit of the lexical-decision model, shipped with the package.

```{r}
draws <- readRDS(system.file("extdata", "lexdec_draws.rds", package = "depictr"))
posterior_plot(draws[c("conditionunrelated", "modalityauditory",
                       "word_frequency")],
               labels = c(conditionunrelated = "condition",
                          modalityauditory = "modality",
                          word_frequency = "word frequency"),
               title = "Lexical-decision fixed effects (ms)")
```

## The shared spine: `tidy_estimates()`

Most of the model functions rest on `tidy_estimates()`, which turns a model, or
a data frame of pre-computed estimates, into one standard table. Because the
plotting functions also accept that table, estimates from any source (Bayesian
posteriors, bootstrap intervals, or figures taken from a paper) can be supplied
directly.

```{r}
tidy_estimates(fit)
```

## A consistent, accessible look

`theme_depictr()`, `depictr_palette()` and `scale_colour_depictr()` style your
own plots too:

```{r}
library(ggplot2)
ggplot(crop_yield, aes(fertiliser, yield, colour = treatment)) +
  geom_point(alpha = 0.7) +
  scale_colour_depictr() +
  theme_depictr()
```

The qualitative palette is based on the Okabe-Ito set [@okabe2008], which stays
distinguishable under the common forms of colour-vision deficiency; sequential
and diverging variants are available too. Preview them with:

```{r, fig.height = 5}
palette_preview(type = "all")
```

`palette_preview()` can also *simulate* a colour-vision deficiency, so you can
check a palette as a deuteranope (red-green) would see it:

```{r, fig.height = 2.6}
palette_preview(cvd = "deutan")
```

Set the look once for a whole script with `depictr_options()` (base size, base
family, brand and accent colours, or a custom palette), instead of passing the
same arguments to every call. Called with no arguments it reports the current
settings:

```{r}
depictr_options()
```

## Where to next

The remaining articles go into each area in turn.
`vignette("exploring-data")` covers distributions, categories, bivariate plots,
scatter-plot matrices, correlations, missingness, outliers, summary tables and
the new estimation plots.
`vignette("model-estimates")` is the flagship: forest plots, model comparison,
predicted values, interactions, random effects, optimiser checks and the
frequentist-over-Bayesian-posterior overlay.
`vignette("diagnostics-and-uncertainty")` covers residuals, GLM-appropriate
binned residuals, the classification suite (ROC, PR, gains, lift, calibration,
thresholds) on an imbalanced outcome, and power curves.
Two further articles, `vignette("multivariate-and-survival")` and
`vignette("time-series")`, cover the remaining methods.

## References
