---
title: "Diagnostics, classification and uncertainty"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Diagnostics, classification and uncertainty}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: ../inst/REFERENCES.bib
csl: apa.csl
link-citations: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      fig.width = 7, fig.height = 4.5, dpi = 96)
library(depictr)
has_simr <- requireNamespace("simr", quietly = TRUE)
```

## Regression diagnostics

`residual_diagnostics_plot()` assembles the classic panel; `influence_plot()`
and `qq_plot()` zoom in on particular checks. We fit a linear model to the
crop-yield trial.

```{r}
fit <- lm(yield ~ rainfall + fertiliser + soil_ph, data = crop_yield)
```

```{r, fig.height = 6}
residual_diagnostics_plot(fit, title = "Crop-yield model")
```

```{r}
influence_plot(fit)
```

The bubble area is Cook's distance, the standard measure of an observation's
influence on the fitted coefficients [@cook1977].

`vif_plot()` checks for multicollinearity among the predictors:

```{r, fig.height = 3.2}
vif_plot(fit)
```

## Diagnostics for a generalised linear model

For a binary `glm` the raw residuals take only a few values, so a plain
residual-versus-fitted plot is hard to read. `binned_residual_plot()` instead
splits the data into equal-count bins of fitted values and plots the mean
residual per bin against a +/- 2 standard-error band [@gelman2007]: most points
should sit inside the band, scattered around zero.

We model the *rare* adverse-event outcome from the clinical trial (about a 10%
base rate) on the baseline biomarker, age and treatment arm.

```{r}
gfit <- glm(adverse_event ~ biomarker + age + arm,
            data = clinical_trial, family = binomial)
```

```{r}
binned_residual_plot(gfit, title = "Binned residuals: adverse-event model")
```

`residual_diagnostics_plot()` recognises a `glm` and switches to
GLM-appropriate panels (binned residuals and a Q-Q plot of randomized quantile
residuals, which are standard-normal under a correct model regardless of the
response distribution):

```{r, fig.height = 6}
residual_diagnostics_plot(gfit, title = "Adverse-event model")
```

## Classification on a genuinely imbalanced outcome

The adverse-event outcome is imbalanced, which is exactly when the
precision-recall, gains and lift charts earn their keep. depictr's classification
plots each read a binomial `glm` directly, a pair of vectors or, to compare
several models, a named list of models.

The ROC curve reports the AUC. Passing `youden = TRUE` marks the Youden's J
operating point (the threshold maximising sensitivity + specificity - 1):

```{r, fig.width = 5, fig.height = 5}
roc_curve_plot(gfit, youden = TRUE)
```

When the positive class is rare the precision-recall curve is more informative
than the ROC curve, because it ignores the many true negatives. The baseline is
the positive prevalence; `f1 = TRUE` marks the maximum-F1 operating point:

```{r, fig.width = 5, fig.height = 5}
pr_curve_plot(gfit, f1 = TRUE)
```

### Comparing two models in one figure

A *named list* overlays one colour-coded curve per model, with a per-curve AUC
(or average precision) in the legend. Here the full model is compared with a
biomarker-only model.

```{r}
reduced <- glm(adverse_event ~ biomarker, data = clinical_trial,
               family = binomial)
models <- list(Full = gfit, `Biomarker only` = reduced)
```

Because an ROC curve hugs the top-left, the bottom-right corner is always free,
so `legend_inside = TRUE` tucks the per-model legend there and reclaims the
right-hand margin:

```{r, fig.width = 5, fig.height = 5}
roc_curve_plot(models, youden = TRUE, legend_inside = TRUE)
```

```{r, fig.width = 5, fig.height = 5}
pr_curve_plot(models, f1 = TRUE)
```

For ranking and targeting tasks, the cumulative gains and lift charts show how
many positive cases are captured as more of the score-ordered population is
targeted, again overlaying both models:

```{r, fig.width = 5, fig.height = 5}
gain_plot(models, legend_inside = TRUE)
```

```{r, fig.height = 3.6}
lift_plot(models, legend_inside = TRUE)
```

### Choosing an operating point

`threshold_plot()` sweeps the decision threshold across the full range of scores
and plots each metric as it trades off, marking the Youden and maximum-F1
optimal thresholds with dashed lines. It is the natural companion to the ROC and
PR curves for actually *choosing* a cut-off.

```{r}
tp <- threshold_plot(gfit, title = "Metrics across the decision threshold")
tp
attr(tp, "thresholds")   # the Youden and max-F1 thresholds
```

### Calibration

`calibration_plot()` checks whether predicted probabilities match observed
frequencies. Each bin's observed rate carries a Wilson binomial confidence
interval, so bins backed by few observations (common in the sparse upper tail
when the outcome is rare) are not over-interpreted.

```{r, fig.width = 5, fig.height = 5}
calibration_plot(gfit, bins = 6)
```

A confusion-matrix heatmap completes the picture. Passing `threshold = "youden"`
reuses the same operating point the ROC curve marks, so the two agree:

```{r, fig.width = 5, fig.height = 4.5}
confusion_matrix_plot(gfit, threshold = "youden", normalise = "row")
```

## Uncertainty

`posterior_plot()` summarises draws (posterior, bootstrap, simulation) as a
distribution per parameter. Here are the real posterior draws from a Bayesian fit
of the lexical-decision model, shown as point-and-interval forests.

```{r}
draws <- readRDS(system.file("extdata", "lexdec_draws.rds", package = "depictr"))
posterior_plot(draws[c("conditionunrelated", "modalityauditory",
                       "word_frequency")],
               labels = c(conditionunrelated = "condition",
                          modalityauditory = "modality",
                          word_frequency = "word frequency"),
               style = "interval", title = "Posterior estimates (ms)")
```

## Power curves

`power_curve_plot()` reads a `simr::powerCurve()` object or a tidy data frame, so
a slow power simulation does not have to be re-run to redraw it. The package
ships a real `powerCurve` object from a `simr` analysis of the lexical-decision
design (power for the condition effect across numbers of participants); we read
it straight from disk.

```{r, eval = has_simr}
pc <- readRDS(system.file("extdata", "powercurve_lexdec.rds", package = "depictr"))
power_curve_plot(pc, x_lab = "Number of participants",
                 title = "Power for the condition effect")
```

```{r, eval = !has_simr, echo = FALSE}
# Summarising a powerCurve object needs the 'simr' package. Without it we draw
# the same curve from a tidy data frame, which power_curve_plot() also accepts.
pc_df <- data.frame(
  nlevels = seq(10, 70, by = 10),
  mean  = c(0.16, 0.31, 0.48, 0.63, 0.78, 0.88, 0.94),
  lower = c(0.09, 0.23, 0.39, 0.54, 0.70, 0.82, 0.89),
  upper = c(0.25, 0.40, 0.57, 0.71, 0.85, 0.93, 0.97)
)
```

```{r, eval = !has_simr, fig.height = 4.5}
power_curve_plot(pc_df, x_lab = "Number of participants",
                 title = "Power for the condition effect")
```

## Composing and saving

Combine any of these with `arrange_plots()` and save with `save_plot()`:

```{r, fig.height = 3.2}
arrange_plots(
  qq_plot(fit), influence_plot(fit),
  ncol = 2, title = "Diagnostics", tag_levels = "A"
)
```

## References
