---
title: "Diagnostics, classification and uncertainty"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Diagnostics, classification and uncertainty}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      fig.width = 7, fig.height = 4.5, dpi = 96)
library(depictr)
```

## Regression diagnostics

`residual_diagnostics_plot()` assembles the classic panel; `influence_plot()`
and `qq_plot()` zoom in on particular checks.

```{r}
fit <- lm(yield ~ rainfall + fertilizer + soil_ph, data = crop_yield)
```

```{r, fig.height = 6}
residual_diagnostics_plot(fit, title = "Crop-yield model")
```

```{r}
influence_plot(fit)
```

`vif_plot()` checks for multicollinearity among the predictors:

```{r, fig.height = 3.2}
vif_plot(fit)
```

## Classification

For binary classifiers, depictr provides an ROC curve (with AUC), a calibration
plot and a confusion-matrix heatmap. Each reads a binomial `glm` directly, or a
pair of vectors.

```{r}
gfit <- glm(accuracy ~ word_frequency + RT + condition,
            data = lexical_decision, family = binomial)
```

```{r, fig.width = 5, fig.height = 5}
roc_curve_plot(gfit)
```

When the positive class is rare, the precision-recall curve is more informative
than the ROC curve:

```{r, fig.width = 5, fig.height = 5}
pr_curve_plot(gfit)
```

For ranking and targeting tasks, the cumulative gains and lift charts show how
many positive cases are captured as more of the score-ordered population is
targeted:

```{r, fig.width = 5, fig.height = 5}
gain_plot(gfit)
```

```{r, fig.height = 3.6}
lift_plot(gfit)
```

```{r, fig.width = 5, fig.height = 5}
calibration_plot(gfit, bins = 8)
```

```{r, fig.width = 5, fig.height = 4.5}
confusion_matrix_plot(gfit, threshold = 0.5, normalize = "row")
```

## Uncertainty

`posterior_plot()` summarises draws (posterior, bootstrap, simulation) as a
point with nested intervals.

```{r}
set.seed(2)
draws <- data.frame(
  stress = rnorm(2000, -0.45, 0.08),
  sleep = rnorm(2000, 0.25, 0.06),
  exercise = rnorm(2000, 0.15, 0.05)
)
posterior_plot(draws, title = "Posterior estimates")
```

## Power curves

`power_curve_plot()` reads a `simr::powerCurve()` object or a tidy data frame,
so a slow power simulation does not have to be re-run to redraw it.

```{r}
pc <- data.frame(nlevels = seq(10, 70, by = 10),
                 mean  = c(0.16, 0.31, 0.48, 0.63, 0.78, 0.88, 0.94),
                 lower = c(0.09, 0.23, 0.39, 0.54, 0.70, 0.82, 0.89),
                 upper = c(0.25, 0.40, 0.57, 0.71, 0.85, 0.93, 0.97))
power_curve_plot(pc, x_lab = "Number of participants",
                 title = "Power for the condition effect")
```

## Composing and saving

Combine any of these with `arrange_plots()` and save with `save_plot()`:

```{r, fig.height = 3.2}
arrange_plots(
  qq_plot(fit), influence_plot(fit),
  ncol = 2, title = "Diagnostics", tag_levels = "A"
)
```
