---
title: "Exploring data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exploring data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      fig.width = 7, fig.height = 4.5, dpi = 96)
library(depictr)
```

depictr provides a coherent set of exploratory plots, a descriptive table and
two estimation plots, all sharing the package theme and palette. Column names
can be given quoted or unquoted. The examples use the bundled `lexical_decision`,
`wellbeing_survey` and `crop_yield` datasets.

## One variable

`explore_distribution()` for a numeric variable, `explore_categorical()` for a
categorical one.

```{r}
explore_distribution(lexical_decision, RT, group = condition, type = "density")
```

```{r}
explore_categorical(wellbeing_survey, education, group = region,
                    proportion = TRUE, position = "dodge")
```

## Two variables, any types

`explore_bivariate()` selects the appropriate plot automatically: a scatter
plot for two numeric variables, box plots for a numeric variable against a
categorical one, and a filled bar chart for two categorical variables.

```{r}
explore_bivariate(lexical_decision, condition, RT)
```

For a focused scatter with a fitted trend, use `scatter_trend()`. The crop-yield
trial has a real fertiliser-by-treatment interaction, visible here as two
diverging slopes.

```{r}
scatter_trend(crop_yield, fertilizer, yield, group = treatment)
```

## Many variables at once

`explore_pairs()` is a scatter-plot matrix; `correlation_heatmap()` condenses
the same information into a single coloured grid.

```{r, fig.height = 6}
explore_pairs(crop_yield, cols = c("rainfall", "fertilizer", "soil_ph", "yield"))
```

```{r, fig.height = 5}
correlation_heatmap(wellbeing_survey)
```

## Distributions across groups

`raincloud_plot()` shows the density, the box summary and the raw points
together, and `group_comparison_plot()` adds the group means with confidence
intervals over the raw data. By showing the estimate alongside its uncertainty,
the latter conveys whether the groups differ more faithfully than a bar chart.

```{r}
raincloud_plot(lexical_decision, RT, group = condition)
```

```{r}
group_comparison_plot(lexical_decision, RT, condition)
```

## Estimation plots: effect size, not just a p-value

An *estimation plot* puts the effect size and its uncertainty at the centre of
the comparison. `estimation_plot()` draws the classic Gardner-Altman two-group
plot: the raw data with group means on top, and the **mean difference** with a
bootstrap confidence interval beneath, aligned so a difference of zero sits under
the reference group's mean. With two groups it also annotates a standardised
effect size (Hedges' *g* by default).

```{r, fig.height = 5}
set.seed(1)
estimation_plot(lexical_decision, RT, condition,
                title = "RT difference: unrelated vs. related priming")
```

`group_comparison_plot(differences = TRUE)` is the same idea reached from the
group-means plot: it appends the difference panel, turning a means comparison
into a full estimation plot. With more than two groups every other group is
compared against a chosen reference (a Cumming plot).

```{r, fig.height = 5}
set.seed(1)
group_comparison_plot(crop_yield, yield, treatment, differences = TRUE,
                      title = "Yield difference: enhanced vs. standard")
```

## Data quality: outliers and missingness

```{r}
outlier_plot(crop_yield, yield)
```

`wellbeing_survey` has *informative* missingness -- income is missing more often
at higher stress -- so the missingness map is worth a look before modelling.

```{r, fig.height = 5}
missingness_map(wellbeing_survey)
```

## A descriptive summary table

`summary_table()` builds a "Table 1": mean (SD) for numeric variables, counts
and percentages for categorical ones, optionally by group. The first row always
reports the sample size (`N`), and any variable with missing values gets a
`Missing, n (%)` row -- both visible here for the wellbeing survey. It returns a
plain data frame, ready for `knitr::kable()`.

```{r}
tab <- summary_table(
  wellbeing_survey,
  vars = c("life_satisfaction", "income", "stress", "education"),
  group = "region"
)
knitr::kable(tab)
```
