---
title: "Exploring data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exploring data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      fig.width = 7, fig.height = 4.5, dpi = 96)
library(statviz)
```

statviz provides a coherent set of exploratory plots and a descriptive table,
all sharing the package theme and palette. Column names can be given quoted or
unquoted.

## One variable

`explore_distribution()` for a numeric variable, `explore_categorical()` for a
categorical one.

```{r}
explore_distribution(lexical_decision, RT, group = condition, type = "density")
```

```{r}
explore_categorical(wellbeing_survey, education, group = region,
                    proportion = TRUE, position = "dodge")
```

## Two variables, any types

`explore_bivariate()` picks the appropriate plot automatically -- scatter for
two numerics, boxplots for numeric-by-categorical, a filled bar chart for two
categoricals.

```{r}
explore_bivariate(lexical_decision, condition, RT)
```

For a focused scatter with a fitted trend, use `scatter_trend()`:

```{r}
scatter_trend(crop_yield, fertilizer, yield, group = treatment)
```

## Many variables at once

`explore_pairs()` is a scatter-plot matrix; `correlation_heatmap()` condenses
the same information into a single coloured grid.

```{r, fig.height = 6}
explore_pairs(crop_yield, cols = c("rainfall", "fertilizer", "soil_ph", "yield"))
```

```{r, fig.height = 5}
correlation_heatmap(wellbeing_survey)
```

## Distributions across groups

`raincloud_plot()` shows the density, the box summary and the raw points
together, and `group_comparison_plot()` adds the group means with confidence
intervals over the raw data -- a clearer answer to "do these groups differ?"
than a bar chart.

```{r}
raincloud_plot(lexical_decision, RT, group = condition)
```

```{r}
group_comparison_plot(lexical_decision, RT, condition)
```

## Data quality: outliers and missingness

```{r}
outlier_plot(crop_yield, yield)
```

```{r, fig.height = 5}
missingness_map(wellbeing_survey)
```

## A descriptive summary table

`summary_table()` builds a "Table 1": mean (SD) for numeric variables, counts
and percentages for categorical ones, optionally by group. It returns a plain
data frame, ready for `knitr::kable()`.

```{r}
tab <- summary_table(wellbeing_survey,
                     vars = c("life_satisfaction", "stress", "education"),
                     group = "region")
knitr::kable(tab)
```
