---
title: "Visualising model estimates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualising model estimates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: ../inst/REFERENCES.bib
csl: apa.csl
link-citations: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      fig.width = 7, fig.height = 4.5, dpi = 96)
library(depictr)
has_lme4   <- requireNamespace("lme4", quietly = TRUE)
has_lmer   <- requireNamespace("lmerTest", quietly = TRUE)
has_ggdist <- requireNamespace("ggdist", quietly = TRUE)
```

This is the flagship article. It works through depictr's model-result plots
(coefficients, model comparison, predicted values, interactions, random effects
and goodness-of-fit) and then reaches the capability the package is named for:
showing a frequentist estimate and a full Bayesian posterior for the same model
on one figure.

The running example is `lexical_decision`, a counterbalanced, crossed
lexical-decision experiment (24 participants, 40 items, 960 trials). We model
reaction time on correct trials as a function of priming `condition`
(related/unrelated), presentation `modality` (visual/auditory) and item
`word_frequency`, with crossed random intercepts for participant and item. The
design is counterbalanced, so neither fixed factor is collinear with the item
random effect, a clean fit for a mixed model.

```{r fit, eval = has_lmer}
correct <- subset(lexical_decision, accuracy == 1)
fit <- lmerTest::lmer(
  RT ~ condition + modality + word_frequency +
    (1 | participant) + (1 | item),
  data = correct
)
```

```{r fit-fallback, eval = !has_lmer, echo = FALSE}
# Without lmerTest/lme4 the live fit is skipped; the remaining chunks that need
# `fit` are guarded by the same condition so the article still knits.
```

## Coefficient (forest) plots

`coefficient_plot()` reads the fitted mixed model directly (through
`tidy_estimates()`) and draws a horizontal point-and-interval ('forest') plot.
Unrelated primes and auditory presentation both slow responses, while more
frequent words speed them up.

```{r coef, eval = has_lmer}
coefficient_plot(
  fit, order = "ascending",
  labels = c(conditionunrelated = "Unrelated priming",
             modalityauditory = "Auditory modality",
             word_frequency = "Word frequency (Zipf)"),
  title = "Predictors of lexical-decision RT (ms)"
)
```

Coefficient names are tidied to the effect (variable) name automatically
(`conditionunrelated` becomes `condition`, `word_frequency` becomes `word
frequency`); pass `labels` to override any of them. The slopes also sit on quite
different scales, so `standardise = TRUE` rescales each
coefficient by its predictor's standard deviation, putting them on one
comparable axis:

```{r coef-std, eval = has_lmer}
coefficient_plot(fit, standardise = TRUE, order = "ascending",
                 title = "Standardised predictors of RT")
```

To keep the raw units instead, including a large intercept that would
otherwise squash the slopes, `facet = TRUE` gives each term its own
free-scaled panel (the layout the frequentist-vs-Bayesian comparison below uses
by default).

## Comparing models

`compare_models()` overlays the estimates from several models so you can see how
a coefficient moves as the specification changes. Here a reduced model (dropping
`modality`) is compared with the full model; `model_fit_table()` summarises their
fit.

```{r compare, eval = has_lmer}
reduced <- lmerTest::lmer(
  RT ~ condition + word_frequency + (1 | participant) + (1 | item),
  data = correct
)
compare_models(Reduced = reduced, Full = fit, order = "descending")
```

```{r fit-table, eval = has_lmer}
knitr::kable(model_fit_table(Reduced = reduced, Full = fit))
```

For `glm` models the `R2` column reports McFadden's pseudo-R-squared
[@mcfadden1974] in place of the ordinary coefficient of determination.

## Predicted values and interactions

`effects_plot()` shows what the model predicts as one predictor varies, holding
the others at typical values. It now works on the mixed model: predictions use
the fixed effects only (`re.form = NA`) and the band is built from the
fixed-effect design matrix and `vcov()`.

```{r effects, eval = has_lmer}
effects_plot(fit, "word_frequency",
             title = "Predicted RT across word frequency")
```

`interaction_plot()` shows how a relationship changes across a second predictor.
The lexical-decision model is additive, so for a *genuine* interaction we switch
to `crop_yield`, whose data-generating process contains a real
fertiliser-by-treatment effect: fertiliser raises yield far more under the
`enhanced` treatment than under `standard`, so the slopes diverge.

```{r interaction}
crop_fit <- lm(yield ~ fertiliser * treatment + rainfall, data = crop_yield)
interaction_plot(crop_fit, "fertiliser", "treatment",
                 title = "Fertiliser x treatment interaction")
```

## Frequentist and Bayesian estimates together

This is the package's namesake capability. Given the frequentist fit and a
set of Bayesian posterior draws, `frequentist_bayesian_plot()` draws the
*full posterior distribution* for each term (a 'ggdist' half-eye) and overlays
the frequentist point and confidence interval at the same position. The entire
shape of the posterior appears next to the frequentist estimate, rather than a
point and two limits alone, with the two sources in the two leading
colourblind-safe palette colours.

The draws here are the real fixed-effect posterior from a `brms` fit of the
same model (1000 draws x 4 parameters), shipped with the package so the slow
MCMC need not be re-run. Terms are matched by canonical label, so the `brms`
parameter names line up with the frequentist ones automatically.

```{r fbp, eval = has_lmer && has_ggdist, fig.height = 4}
draws <- readRDS(system.file("extdata", "lexdec_draws.rds", package = "depictr"))

frequentist_bayesian_plot(
  fit, draws,
  intercept = FALSE,
  note_frequentist_no_prior = TRUE,
  title = "Frequentist estimate over the full Bayesian posterior"
)
```

The frequentist confidence interval and the bulk of the Bayesian posterior land
in the same place (reassuring agreement between the two paradigms), but only
the posterior shows the density, the skew and the mass on either side of zero.

If `ggdist` is unavailable the function falls back to a point-and-interval forest
plot of the two sources, and when the Bayesian side is supplied as a *summary*
table instead of draws (for instance the `Estimate`/`Q2.5`/`Q97.5` of
`brms::fixef()`) it draws the familiar two-source forest plot.

## Posterior distributions on their own

`posterior_plot()` summarises any draws (posterior, bootstrap or simulation) as
a distribution per parameter. With `style = "halfeye"` it shows the density slab
and a point-and-interval; a region of practical equivalence (ROPE) can be shaded
and each parameter annotated with its probability of direction (the posterior
mass on its majority side of the reference line).

```{r posterior, eval = has_ggdist, fig.height = 4}
draws <- readRDS(system.file("extdata", "lexdec_draws.rds", package = "depictr"))
slopes <- draws[c("conditionunrelated", "modalityauditory", "word_frequency")]

posterior_plot(
  slopes, style = "halfeye", rope = c(-5, 5), pd = TRUE,
  labels = c(conditionunrelated = "condition",
             modalityauditory = "modality",
             word_frequency = "word frequency"),
  title = "Fixed-effect posteriors (ms), with ROPE and pd"
)
```

The probability of direction for `word_frequency` and `conditionunrelated` is
effectively 100%: the posterior sits almost entirely on one side of zero.

## Random effects

`random_effects_plot()` draws a caterpillar plot of the conditional modes
('BLUPs'). Reading the fitted model directly, it shows the by-item and
by-participant departures from the average, sorted, with their uncertainty, the
usual way to spot unusual groups.

```{r ranef, eval = has_lme4, fig.height = 6}
random_effects_plot(fit, title = "By-group departures (random intercepts)")
```

## Optimiser checks

A mixed-model fit should be stable across optimisers. `lme4::allFit()` refits the
model with every available optimiser; `optimizer_fixef_plot()` then shows the
fixed effects side by side, one panel per term. Tight clusters mean the fit has
settled; scatter would signal a fragile solution.

The package ships the `allFit()` summary for this model, so we can plot it
without re-running the (slow) refits. The plot accepts a tidy data frame of
optimiser-by-term values, which we read straight off the stored summary.

```{r optim, fig.height = 4.5}
af <- readRDS(system.file("extdata", "allfit_lexdec.rds", package = "depictr"))
fx <- af$fixef    # optimisers x fixed effects
opt_long <- data.frame(
  optimizer = rep(rownames(fx), times = ncol(fx)),
  term      = rep(colnames(fx), each = nrow(fx)),
  value     = as.vector(fx)
)

optimizer_fixef_plot(
  opt_long, title = "Fixed effects across optimisers",
  labels = c(conditionunrelated = "condition",
             modalityauditory = "modality",
             word_frequency = "word frequency")
)
```

Every optimiser lands on the same estimate for each term (the points coincide
within each panel), so this fit is stable.

## A one-figure model report

`model_report()` composes several views (coefficients, the effect of a focal
predictor, residuals against fitted values and a Q-Q plot, with a fit-statistics
subtitle) into a single figure for a rapid review or a report appendix. It works
on `lm`/`glm` models; here we use the crop-yield model.

```{r report, fig.width = 9, fig.height = 7}
full <- lm(yield ~ rainfall + fertiliser + soil_ph + treatment,
           data = crop_yield)
model_report(full, title = "Crop-yield model")
```

## References
