---
title: "Visualising model estimates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualising model estimates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.5,
  dpi = 96
)
library(modelviz)
```

This vignette covers the model-result plots: coefficient plots, comparisons of
several models, the frequentist/Bayesian overlay, optimiser checks, power
curves and regression diagnostics.

## Coefficient (forest) plots

`coef_plot()` accepts any model `tidy_estimates()` understands, or a tidy data
frame. Customise the term order, the labels, and the reference line.

```{r}
fit <- lm(life_satisfaction ~ stress + sleep_hours + exercise_days,
          data = wellbeing_survey)

coef_plot(
  fit,
  order = "ascending",
  labels = c(stress = "Stress", sleep_hours = "Sleep (h)",
             exercise_days = "Exercise (days)"),
  title = "Predictors of life satisfaction"
)
```

For odds or risk ratios, set the reference line to 1 and label the axis
accordingly.

```{r}
gfit <- glm(accuracy ~ word_frequency + condition, data = lexical_decision,
            family = binomial)

coef_plot(gfit, reference_line = 0, x_lab = "Log-odds estimate",
          title = "Log-odds of a correct response")
```

## Comparing several models

`compare_estimates_plot()` overlays the estimates from two or more sources,
one colour per source. Name the arguments to label the sources.

```{r}
m_all   <- lm(yield ~ rainfall + fertilizer + soil_ph, data = crop_yield)
m_std   <- lm(yield ~ rainfall + fertilizer + soil_ph,
              data = crop_yield[crop_yield$treatment == "standard", ])
m_enh   <- lm(yield ~ rainfall + fertilizer + soil_ph,
              data = crop_yield[crop_yield$treatment == "enhanced", ])

compare_estimates_plot(
  `All fields` = m_all, Standard = m_std, Enhanced = m_enh,
  order = "descending", title = "Estimates by management treatment"
)
```

## Frequentist and Bayesian estimates together

`frequentist_bayesian_plot()` is a focused wrapper for the common case of
showing a frequentist and a Bayesian fit side by side, in the familiar red and
blue styling. The Bayesian summary is usually supplied as a tidy data frame
(for example from `brms::fixef()` or `posterior::summarise_draws()`); here we
build a stand-in so the example is self-contained.

```{r}
freq  <- lm(life_satisfaction ~ stress + sleep_hours + exercise_days,
            data = wellbeing_survey)

bayes <- tidy_estimates(freq)
bayes$estimate  <- bayes$estimate * 0.92          # pretend posterior means
bayes$conf.low  <- bayes$conf.low * 0.92
bayes$conf.high <- bayes$conf.high * 0.92

frequentist_bayesian_plot(
  freq, bayes,
  title = "Frequentist and Bayesian estimates",
  note_frequentist_no_prior = TRUE
)
```

## Checking optimiser stability

`optimizer_fixef_plot()` shows how the fixed-effect estimates of a mixed model
vary across the optimisers tried by `lme4::allFit()`. Tight clusters indicate
agreement; scatter flags a fragile fit. Without lme4 you can build the input
frame directly:

```{r}
opt <- expand.grid(
  optimizer = c("bobyqa", "Nelder_Mead", "nlminbwrap", "nmkbw"),
  term = c("(Intercept)", "stress", "sleep_hours")
)
set.seed(1)
opt$value <- c(4.5, 4.51, 4.49, 4.50,
               -0.45, -0.44, -0.46, -0.45,
               0.25, 0.24, 0.26, 0.25) + rnorm(12, 0, 0.005)

optimizer_fixef_plot(opt, title = "Fixed effects across optimisers")
```

## Power analysis curves

`power_curve_plot()` plots power against sample size with a target line. It
reads a `simr::powerCurve()` object, or a tidy data frame so the (often slow)
simulation does not have to be re-run.

```{r}
pc <- data.frame(
  nlevels = seq(10, 70, by = 10),
  mean  = c(0.16, 0.31, 0.48, 0.63, 0.78, 0.88, 0.94),
  lower = c(0.09, 0.23, 0.39, 0.54, 0.70, 0.82, 0.89),
  upper = c(0.25, 0.40, 0.57, 0.71, 0.85, 0.93, 0.97)
)

power_curve_plot(pc, x_lab = "Number of participants",
                 title = "Power for the condition effect")
```

## Regression diagnostics

`residual_diagnostics_plot()` assembles the classic diagnostic plots into one
panel.

```{r, fig.height = 6}
residual_diagnostics_plot(m_all, title = "Diagnostics: crop-yield model")
```
