---
title: "Visualising model estimates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualising model estimates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      fig.width = 7, fig.height = 4.5, dpi = 96)
library(statviz)
```

This vignette covers the model-result plots: coefficients, model comparison,
predicted values, interactions, random effects and goodness-of-fit.

## Coefficient (forest) plots

`coefficient_plot()` accepts any model `tidy_estimates()` understands, or a tidy
data frame.

```{r}
fit <- lm(life_satisfaction ~ stress + sleep_hours + exercise_days,
          data = wellbeing_survey)
coefficient_plot(fit, order = "ascending",
                 labels = c(stress = "Stress", sleep_hours = "Sleep (h)",
                            exercise_days = "Exercise (days)"),
                 title = "Predictors of life satisfaction")
```

## Comparing models

`compare_models()` overlays the estimates from several models, and
`model_fit_table()` summarises their fit.

```{r}
m1 <- lm(yield ~ rainfall + fertilizer, data = crop_yield)
m2 <- lm(yield ~ rainfall + fertilizer + soil_ph, data = crop_yield)
m3 <- lm(yield ~ rainfall + fertilizer + soil_ph + treatment, data = crop_yield)

compare_models(small = m1, medium = m2, full = m3, order = "descending")
```

```{r}
knitr::kable(model_fit_table(small = m1, medium = m2, full = m3))
```

## Frequentist and Bayesian estimates together

```{r}
freq  <- lm(life_satisfaction ~ stress + sleep_hours + exercise_days,
            data = wellbeing_survey)
bayes <- tidy_estimates(freq)
bayes$estimate  <- bayes$estimate * 0.92      # stand-in posterior means
bayes$conf.low  <- bayes$conf.low * 0.92
bayes$conf.high <- bayes$conf.high * 0.92

frequentist_bayesian_plot(freq, bayes, note_frequentist_no_prior = TRUE,
                          title = "Frequentist and Bayesian estimates")
```

## Predicted values and interactions

`effects_plot()` shows what the model predicts as one predictor varies;
`interaction_plot()` shows how that relationship changes across a second
predictor.

```{r}
fit2 <- lm(yield ~ fertilizer * treatment + rainfall, data = crop_yield)
effects_plot(fit2, "rainfall")
```

```{r}
interaction_plot(fit2, "fertilizer", "treatment")
```

For a binomial `glm`, predictions are shown on the probability scale:

```{r}
gfit <- glm(accuracy ~ word_frequency + condition, data = lexical_decision,
            family = binomial)
effects_plot(gfit, "word_frequency")
```

## Random effects

`random_effects_plot()` draws a caterpillar plot. It reads an 'lme4' model
directly, or a data frame of by-group estimates:

```{r}
re <- data.frame(level = paste0("site_", 1:12),
                 estimate = sort(rnorm(12, 0, 0.5)),
                 std.error = runif(12, 0.1, 0.3))
random_effects_plot(re, title = "By-site departures")
```

## Optimiser checks

When a mixed model is fitted by several optimisers (via `lme4::allFit()`),
`optimizer_fixef_plot()` shows whether they agree. Without 'lme4' you can build
the input directly:

```{r}
opt <- expand.grid(optimizer = c("bobyqa", "Nelder_Mead", "nlminbwrap"),
                   term = c("(Intercept)", "stress", "sleep_hours"))
opt$value <- c(4.5, 4.51, 4.49, -0.45, -0.44, -0.46, 0.25, 0.24, 0.26)
optimizer_fixef_plot(opt)
```

## A one-figure model report

`model_report()` composes several of these views -- coefficients, the effect of
a focal predictor, residuals-vs-fitted and a Q-Q plot, with a fit-statistics
subtitle -- into a single figure for a quick review or a report appendix.

```{r, fig.width = 9, fig.height = 7}
full <- lm(yield ~ rainfall + fertilizer + soil_ph + treatment,
           data = crop_yield)
model_report(full, title = "Crop-yield model")
```
