---
title: "Getting started with modelviz"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with modelviz}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.5,
  dpi = 96
)
```

## Overview

**modelviz** turns the output of common statistical analyses into clear,
publication-ready graphics, and adds a family of exploratory plots for the data
behind those analyses. Every function returns a `ggplot2` object (or, for the
diagnostics panel, a `patchwork` object), so you can keep customising with the
usual `+` syntax.

The package groups its functions into three families:

* **Model-result plots** -- `coef_plot()`, `compare_estimates_plot()`,
  `frequentist_bayesian_plot()`, `optimizer_fixef_plot()`,
  `power_curve_plot()` and `residual_diagnostics_plot()`.
* **Data-exploration plots** -- `distribution_plot()`, `scatter_trend_plot()`,
  `correlation_matrix_plot()` and `missingness_plot()`.
* **Shared building blocks** -- `tidy_estimates()`, `theme_modelviz()`,
  `modelviz_palette()` and `format_terms()`.

```{r}
library(modelviz)
```

The package ships with three reproducibly simulated datasets used throughout
the documentation: `lexical_decision`, `wellbeing_survey` and `crop_yield`.

```{r}
head(crop_yield)
```

## A first plot

Fit a model and visualise its coefficients as a forest plot. The intercept is
dropped by default, terms keep their model order unless you ask otherwise, and
a dashed reference line marks zero.

```{r}
fit <- lm(yield ~ rainfall + fertilizer + soil_ph + treatment,
          data = crop_yield)

coef_plot(fit, order = "descending", title = "Drivers of crop yield")
```

## The shared currency: `tidy_estimates()`

Most model-result functions are built on `tidy_estimates()`, which turns a
model -- or a data frame of pre-computed estimates -- into one standard table.

```{r}
tidy_estimates(fit)
```

Because the plotting functions also accept that table directly, you can plug in
estimates from sources modelviz does not natively understand (Bayesian
posteriors, bootstrapped intervals, results copied from a paper) just by
shaping them into a data frame with `term`, `estimate`, `conf.low` and
`conf.high` columns.

## A consistent look

`theme_modelviz()`, `modelviz_palette()` and `scale_colour_modelviz()` give
every plot -- including your own -- a consistent appearance.

```{r}
library(ggplot2)

ggplot(crop_yield, aes(fertilizer, yield, colour = treatment)) +
  geom_point(alpha = 0.7) +
  scale_colour_modelviz() +
  theme_modelviz() +
  labs(title = "Yield against fertiliser")
```

## Where to next

* `vignette("model-estimates", package = "modelviz")` -- comparing models,
  frequentist vs. Bayesian estimates, optimiser checks, power curves and
  regression diagnostics.
* `vignette("data-exploration", package = "modelviz")` -- distributions,
  correlations, missing-data maps and scatter plots with trends.
