---
title: "Multivariate analysis and survival"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multivariate analysis and survival}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      fig.width = 7, fig.height = 4.5, dpi = 96)
library(statviz)
```

Beyond regression, statviz covers two staples of applied data analysis:
principal component analysis and survival curves.

## Principal component analysis

`pca_plot()` runs a PCA on the numeric columns of a data frame and draws a
biplot -- the observations projected onto two components, with the variable
loadings as arrows. `scree_plot()` shows how much variance each component
explains.

```{r}
num <- c("rainfall", "fertilizer", "soil_ph", "yield")
pca_plot(crop_yield, cols = num, group = "treatment",
         title = "Crop-yield PCA")
```

```{r, fig.height = 4}
scree_plot(crop_yield, cols = num)
```

Both functions also accept a ready-made `prcomp()` object, so you can analyse
once and plot several views:

```{r}
pc <- prcomp(crop_yield[num], scale. = TRUE)
pca_plot(pc, components = c(1, 3))
```

## Survival curves

`survival_plot()` draws Kaplan-Meier curves with stepwise confidence limits and
censoring marks. The estimate is computed in base R, so no modelling package is
needed; you can pass follow-up times and an event indicator directly, a data
frame, or a `survival::survfit` object.

```{r}
set.seed(1)
n <- 250
group <- sample(c("control", "treated"), n, replace = TRUE)
event_time <- rexp(n, rate = ifelse(group == "treated", 0.05, 0.1))
censor_time <- runif(n, 0, 30)
observed <- pmin(event_time, censor_time)
event <- as.integer(event_time <= censor_time)

survival_plot(observed, event, group = group,
              title = "Survival by treatment")
```
