---
title: "Multivariate analysis and survival"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multivariate analysis and survival}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: ../inst/REFERENCES.bib
csl: apa.csl
link-citations: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      fig.width = 7, fig.height = 4.5, dpi = 96)
library(depictr)
```

Beyond regression, depictr covers three staples of applied data analysis:
principal component analysis, clustering (with quality diagnostics) and survival
curves.

## Principal component analysis

`pca_plot()` runs a PCA on the numeric columns of a data frame and draws a
biplot: the observations projected onto two components, with the variable
loadings as arrows. `scree_plot()` shows how much variance each component
explains.

```{r}
num <- c("rainfall", "fertiliser", "soil_ph", "yield")
pca_plot(crop_yield, cols = num, group = "treatment",
         title = "Crop-yield PCA")
```

```{r, fig.height = 4}
scree_plot(crop_yield, cols = num)
```

Both functions also accept a ready-made `prcomp()` object, so you can analyse
once and plot several views:

```{r}
pc <- prcomp(crop_yield[num], scale. = TRUE)
pca_plot(pc, components = c(1, 3))
```

## Clustering

`cluster_plot()` runs k-means and shows the clusters on the first two principal
components (so it works for any number of variables), with convex hulls and
labelled centroids. `dendrogram_plot()` draws a hierarchical-clustering tree and
can cut it into `k` groups.

```{r}
cluster_plot(crop_yield, cols = num, k = 3, seed = 1,
             title = "Crop-yield clusters")
```

```{r, fig.height = 4}
region_means <- aggregate(
  cbind(stress, sleep_hours, life_satisfaction, age, income) ~ region,
  data = wellbeing_survey, FUN = mean
)
rownames(region_means) <- region_means$region
dendrogram_plot(region_means[-1], k = 2, title = "Regions clustered")
```

## How many clusters? Quality diagnostics

Choosing `k` should not be guesswork. `k_diagnostic()` evaluates a
cluster-quality criterion across a range of `k` and suggests a value, using the
average silhouette width (the default, after Rousseeuw, 1987), the
within-sum-of-squares elbow or the gap statistic (Tibshirani, Walther & Hastie,
2001). See `?k_diagnostic` for the references.

```{r, fig.height = 3.5}
kd <- k_diagnostic(crop_yield, k_range = 2:6, cols = num,
                   method = "silhouette")
kd  # the diagnostic curve, with the suggested k marked
```

The suggested `k` and the underlying table are attached as attributes:

```{r}
attr(kd, "suggested")
knitr::kable(attr(kd, "k_table"), digits = 3)
```

The gap statistic compares the within-cluster dispersion against a null
reference, so unlike the other two criteria it can also support `k = 1`:

```{r, fig.height = 3.5}
k_diagnostic(crop_yield, k_range = 2:6, cols = num, method = "gap")
```

`silhouette_plot()` then shows the quality of an actual clustering, one bar per
observation grouped by cluster, with the average silhouette width per cluster and
overall (the dashed line). Wide positive bars are well-placed observations;
negative bars may belong to a neighbouring cluster.

```{r, fig.height = 5}
cl <- kmeans(scale(crop_yield[num]), centers = attr(kd, "suggested"),
             nstart = 10)$cluster
silhouette_plot(crop_yield, cl, cols = num,
                title = "Silhouette widths by cluster")
```

The same diagnostics apply to the wellbeing survey's numeric profile:

```{r, fig.height = 4}
wb_num <- c("age", "income", "stress", "sleep_hours",
            "exercise_days", "life_satisfaction")
wkd <- k_diagnostic(wellbeing_survey, k_range = 2:6, cols = wb_num,
                    method = "wss")
wcl <- kmeans(scale(na.omit(wellbeing_survey[wb_num])),
              centers = attr(wkd, "suggested"), nstart = 10)$cluster
silhouette_plot(na.omit(wellbeing_survey[wb_num]), wcl,
                title = sprintf("Wellbeing survey, k = %d", attr(wkd, "suggested")))
```

## Survival curves

`survival_plot()` draws Kaplan-Meier curves [@kaplan1958] with stepwise
confidence limits from Greenwood's formula [@greenwood1926] and censoring marks.
The estimate is computed in base R, so no modelling package is needed; you can
pass follow-up times and an event indicator directly, a data frame, or a
`survival::survfit` object.

The `clinical_trial` dataset has two arms with a real survival difference (the
treatment arm has a lower hazard), so the curves separate. Turning on the three
publication annotations gives a survminer-style figure: a number-at-risk
table beneath the curves, dashed guides to each arm's median survival and
the log-rank test of the difference as a subtitle. A survival curve is
monotone-decreasing, so its bottom-left corner is always empty;
`legend_inside = TRUE` puts the group legend there.

```{r, fig.height = 6}
survival_plot(
  clinical_trial$time, clinical_trial$event, group = clinical_trial$arm,
  risk_table = TRUE, median_line = TRUE, logrank = TRUE, legend_inside = TRUE,
  x_lab = "Months", title = "Overall survival by arm"
)
```

The log-rank p-value is tiny and the median survival is clearly longer in the
treatment arm. Because its event times are longer, that arm is more often
right-censored at the 36-month study end, visible as the denser run of censoring
marks.

## References
