---
title: "Getting started with statviz"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with statviz}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      fig.width = 7, fig.height = 4.5, dpi = 96)
```

**statviz** is a single, consistent toolkit of plots that span the whole
analysis workflow -- from a first look at the data, through model estimates and
predictions, to diagnostics, uncertainty and reporting. Every plotting function
returns a `ggplot2` object (or a `patchwork` for composite panels), so you can
keep customising with the usual `+` syntax, and every plot shares one theme, one
palette and one set of label conventions.

```{r}
library(statviz)
```

The package ships with three reproducibly simulated datasets used throughout the
documentation: `lexical_decision`, `wellbeing_survey` and `crop_yield`.

## One example per family

**Explore the data.** Pick the right plot for any pair of variables:

```{r}
explore_bivariate(crop_yield, fertilizer, yield)
```

**Look at model estimates.** Fit a model and draw a forest plot:

```{r}
fit <- lm(yield ~ rainfall + fertilizer + soil_ph + treatment, data = crop_yield)
coefficient_plot(fit, order = "descending", title = "Drivers of crop yield")
```

**Show predictions.** What does the model predict as one predictor varies?

```{r}
effects_plot(fit, "fertilizer")
```

**Check the fit.** A residual-diagnostics panel:

```{r, fig.height = 6}
residual_diagnostics_plot(fit)
```

**Convey uncertainty.** Summarise posterior or simulation draws:

```{r}
set.seed(1)
draws <- data.frame(intercept = rnorm(2000, 2, 0.2),
                    fertilizer = rnorm(2000, 0.012, 0.002),
                    rainfall = rnorm(2000, 0.004, 0.001))
posterior_plot(draws)
```

## The shared spine: `tidy_estimates()`

Most model functions are built on `tidy_estimates()`, which turns a model -- or
a data frame of pre-computed estimates -- into one standard table. Because the
plotting functions also accept that table, you can plug in estimates from any
source (Bayesian posteriors, bootstrap intervals, numbers from a paper).

```{r}
tidy_estimates(fit)
```

## A consistent, reusable look

`theme_statviz()`, `statviz_palette()` and `scale_colour_statviz()` style your
own plots too:

```{r}
library(ggplot2)
ggplot(crop_yield, aes(fertilizer, yield, colour = treatment)) +
  geom_point(alpha = 0.7) +
  scale_colour_statviz() +
  theme_statviz()
```

## Where to next

* `vignette("exploring-data")` -- distributions, categories, bivariate plots,
  scatter-plot matrices, correlations, missingness, outliers and summary tables.
* `vignette("model-estimates")` -- forest plots, model comparison, predicted
  values, interactions, random effects and goodness-of-fit.
* `vignette("diagnostics-and-uncertainty")` -- residuals, influence, Q-Q, ROC,
  calibration, confusion matrices, posteriors and power curves.
