---
title: "Time series"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Time series}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      fig.width = 7, fig.height = 4, dpi = 96)
library(depictr)
```

depictr provides a small, consistent set of time-series plots. The examples use
the bundled `monthly_sales` dataset: two product lines (`indoor` and `outdoor`),
each with six years of monthly observations carrying a trend, a twelve-month
seasonal cycle and noise.

## Plotting one or several series

`timeseries_plot()` accepts a `ts` object, a numeric vector, or a data frame with
time, value and (optionally) group columns. Passing the long-form data frame with
a grouping column draws both series at once, each in its own colour; a
moving-average overlay is one argument away.

```{r}
timeseries_plot(monthly_sales, time = date, value = sales, group = series,
                rolling = 12, title = "Monthly sales by product line",
                y_lab = "Units")
```

For the single-series views (decomposition, autocorrelation, the seasonal plot
and forecasting) we extract one line as a monthly `ts` object of frequency 12.

```{r}
indoor <- subset(monthly_sales, series == "indoor")
indoor <- indoor[order(indoor$date), ]
indoor_ts <- ts(indoor$sales, start = c(2018, 1), frequency = 12)
```

## Decomposition

`decompose_plot()` separates a seasonal series into trend, seasonal and
remainder components. Use `method = "stl"` (the default, loess-based) or
`method = "classical"`.

```{r, fig.height = 6}
decompose_plot(indoor_ts, title = "Indoor sales, decomposed")
```

## Autocorrelation

`acf_plot()` draws the autocorrelation (or, with `type = "partial"`, the partial
autocorrelation) function, with approximate significance bounds. The spikes at
multiples of twelve are the annual seasonality.

```{r, fig.height = 3.4}
acf_plot(indoor_ts)
```

```{r, fig.height = 3.4}
acf_plot(indoor_ts, type = "partial")
```

## The seasonal pattern up close

`seasonal_plot()` draws a seasonal-subseries (cycle) plot: one small panel per
month, with the value traced across successive years and a reference line at each
month's mean. This shows the seasonal shape (differences *between* panels) and
the year-on-year trend within each month (the slope *inside* each panel) at the
same time, something a single overlaid line cannot do.

```{r, fig.height = 3.8}
seasonal_plot(indoor_ts, title = "Indoor sales: monthly subseries")
```

With `style = "season"` every year becomes its own line over the months on a
shared axis, which is handy for spotting an unusual year.

```{r, fig.height = 4}
seasonal_plot(indoor_ts, style = "season",
              title = "Indoor sales: one line per year")
```

## Forecasting

`ts_forecast()` is a lightweight, dependency-free forecaster: it decomposes the
series with STL, extrapolates the recent trend, carries the seasonal pattern
forward, and returns point forecasts with prediction intervals that widen with
the horizon.

```{r}
fc <- ts_forecast(indoor_ts, h = 18, level = 0.9)
head(fc)
```

Passing an integer horizon straight to `timeseries_plot()` overlays that forecast
on the history: the point forecast continues the line and the shaded ribbon shows
the (growing) 90% prediction interval.

```{r, fig.height = 4}
timeseries_plot(indoor_ts, forecast = 18, level = 0.9,
                title = "Indoor sales with an 18-month forecast",
                y_lab = "Units")
```

For a fully specified statistical model, fit it yourself (for example with
`forecast::forecast()`) and pass the resulting `time`/`fit`/`lwr`/`upr` columns
to `timeseries_plot(forecast = )` as a data frame.
