---
title: "Time series"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Time series}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      fig.width = 7, fig.height = 4, dpi = 96)
library(depictr)
```

depictr provides a small, consistent set of time-series plots. The examples use
`AirPassengers`, a classic monthly series that ships with base R.

## Plotting a series

`timeseries_plot()` accepts a `ts` object, a numeric vector, or a data frame
with time, value and (optionally) group columns. A moving-average overlay is one
argument away.

```{r}
timeseries_plot(AirPassengers, rolling = 12,
                title = "Air passengers", y_lab = "Passengers (thousands)")
```

## Autocorrelation

`acf_plot()` draws the autocorrelation (or, with `type = "partial"`, the partial
autocorrelation) function, with approximate significance bounds.

```{r, fig.height = 3.4}
acf_plot(AirPassengers)
```

```{r, fig.height = 3.4}
acf_plot(AirPassengers, type = "partial")
```

## Decomposition

`decompose_plot()` separates a seasonal series into trend, seasonal and
remainder components. Use `method = "stl"` (the default, loess-based) or
`method = "classical"`.

```{r, fig.height = 6}
decompose_plot(AirPassengers, title = "Air passengers, decomposed")
```
